Well-quasi-orderings on word languages

### 2025-10-15 src/appendix.tex
Move some stuff to appendix

@@

>>>>

\begin{proofof}{prefix-wqo-reg-decidable:cor}
	If $L$ is regular, then it is $\MSO$-definable, and there 
	exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
	of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
	exists an \kl{antichain branch} for $L$, we can simply check whether
	the following formula is satisfied:
	\begin{equation*}
		\exists B. 
		B \text{ is a branch } \land
		\forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
		\quad .
	\end{equation*}
	Because the above formula is an $\MSO$-formula over the infinite
	$\Sigma$-branching tree, whether it is satisfied is decidable
	as an easy consequence of the decidability of $\MSO$ over infinite binary
	trees
	\cite[Theorem 1.1]{RAB69}.
\end{proofof}

@@

\begin{proofof}{bounded-language:lem}
	\begin{proof}
		Let $w_1, \dots, w_n$ be such that
		$L \subseteq w_1^* \cdots w_n^*$.
		Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$

		Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
		to $\Sigma^*$. We are interested in the intersection of the image of $w$
		with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
		there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
		Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
		$x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
		$w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
		\InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
		\cup \InfPeriodChain{x} \InfPeriodChain{y}$.

		Now, it may be the case that one cannot simultaneously assume that two
		component of the vector $\vec{k}$ are unbounded. In general, given a set $S
		\subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
		there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
		exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
		$\vec{b}$ on the $S$ components, and the other components are below the
		bound $N_0$. The language of an admissible set $S$ is the set of words
		obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
		($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
		Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.

		Now, admissible languages are ready to be pumped according to
		\cref{pumping-periods:lem}. For every admissible language,
		the size of a word that is not iterated is at most
		$N_0 \times m$ by definition, and we conclude that:
		\begin{equation}
			\label{bounded-language:eq}
			L \subseteq 
			\bigcup_{x,y \in \Sigma^{\leq n}}
			\bigcup_{u \in \Sigma^{\leq m \times N_0}}
			\InfPeriodChain{x} u \InfPeriodChain{y}
			\cup
			\InfPeriodChain{x} u
			\cup
			u \InfPeriodChain{x}
			\quad .
		\end{equation}
	\end{proof}
\end{proofof}

### 2025-10-15 src/infixes.tex
new: moving stuff to appendix
Move some stuff to appendix

\begin{lemma}
    \label{bounded-language:lem}
    Let $L \subseteq \Sigma^*$ be a \kl{bounded language}
    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
    Then, there exists a finite subset $E \subseteq (\Sigma^*)^3$,
    such that:
    \begin{equation*}
        L \subseteq \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}
        \quad .
    \end{equation*}
\end{lemma}
\begin{proof}
    Let $w_1, \dots, w_n$ be such that
    $L \subseteq w_1^* \cdots w_n^*$.
    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$

    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
    to $\Sigma^*$. We are interested in the intersection of the image of $w$
    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.

    Now, it may be the case that one cannot simultaneously assume that two
    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
    $\vec{b}$ on the $S$ components, and the other components are below the
    bound $N_0$. The language of an admissible set $S$ is the set of words
    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.

    Now, admissible languages are ready to be pumped according to
    \cref{pumping-periods:lem}. For every admissible language,
    the size of a word that is not iterated is at most
    $N_0 \times m$ by definition, and we conclude that:
    \begin{equation}
        \label{bounded-language:eq}
        L \subseteq 
        \bigcup_{x,y \in \Sigma^{\leq n}}
        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
        \InfPeriodChain{x} u \InfPeriodChain{y}
        \cup
        \InfPeriodChain{x} u
        \cup
        u \InfPeriodChain{x}
        \quad .
    \end{equation}
\end{proof}

>>>>

\begin{lemma}
    \label{bounded-language:lem}
    \proofref{bounded-language:lem}
    Let $L \subseteq \Sigma^*$ be a \kl{bounded language}
    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
    Then, there exists a finite subset $E \subseteq (\Sigma^*)^3$,
    such that:
    \begin{equation*}
        L \subseteq \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}
        \quad .
    \end{equation*}
\end{lemma}

### 2025-10-15 src/prefixes.tex
Move some stuff to appendix

\begin{corollary}
    \label{prefix-wqo-reg-decidable:cor}
    If $L$ is regular, then the existence of an infinite antichain is decidable.
\end{corollary}
\begin{proof}
    If $L$ is regular, then it is $\MSO$-definable, and there 
    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
    exists an \kl{antichain branch} for $L$, we can simply check whether
    the following formula is satisfied:
    \begin{equation*}
        \exists B. 
        B \text{ is a branch } \land
        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
        \quad .
    \end{equation*}
    Because the above formula is an $\MSO$-formula over the infinite
    $\Sigma$-branching tree, whether it is satisfied is decidable
    as an easy consequence of the decidability of $\MSO$ over infinite binary
    trees
    \cite[Theorem 1.1]{RAB69}.
\end{proof}

>>>>

\begin{corollary}
    \label{prefix-wqo-reg-decidable:cor}
    \proofref{prefix-wqo-reg-decidable:cor}
    If $L$ is regular, then the existence of an infinite antichain is decidable.
\end{corollary}
